\sysfull

### 2017-11-13 paper/coalescing.tex
Work on task coaleascing (not ready yet)

\sys, as the first task-based execution model, dynamically coalesces consecutively executing tasks at run time. When $X$ tasks are coalesced, the $X-1$ tasks do not commit their updates to protected memory locations when reaching a \transition statement. Instead, execution immediately continues from the beginning of the $X$th task. At the end of the $X$th coalesced task, \sys commits the pages that were updated by any of the $X$ coalescing tasks. If consecutive tasks access the same pages of data, the pages accessed by both tasks only need to be committed once.

\textbf{Task Coalescing Problem:} \sys can coalesce an arbitrary number of consecutive tasks. As more tasks coalesce, their collective commit overhead amortizes better, potentially improving performance. As more tasks coalesce, the amount of potentially {\em wasted work} also increases. If power fails during a long sequence of coalesced tasks the lack of an intervening commit requires execution to restart from the {\em first} task in the sequence, failing to preserve the progress of any of the tasks in the long sequence. \sys moderates the risk of coalescing and capitalizes on its benefit by dynamically, \emph{adaptively} determining a coalescing factor that combines many tasks without suffering an excess of wasted work.

The challenge to coalescing tasks is determining how many tasks to coalesce before committing. During execution, an executing task is either a {\em coalesced} task, which does not commit its state before transitioning or a {\em committing} task, which does commit its state before transitioning. \sys monitors a program's execution to adaptively determine $C$, the number of tasks to coalesce before committing; \sys repeatedly executes $C-1$ coalesced tasks,
followed by a committing task.

\textbf{Task Coalescing Strategies:} There are infinite possible strategies for coalescing (infinite amount of $(C_U,C_D)$ pairs). We will consider the three fundamental ones:

\begin{enumerate}
	\item \textbf{$N$ up/$N$ down:} where task coalescing factor increases by $N$ at each successful virtual task execution and decreases by $N$, otherwise;
	\item \textbf{History-driven:} where coalescing will consider history of power failures into deciding on $(C_U,C_D)$ values on the fly;
	\item \textbf{Program profiling-driven:} where coalescing strategy takes into account \emph{expected size} of each task in the application to pick the best coalescing factor.
\end{enumerate}

\subsection{Operation of $N$ up/$N$ Down Coalescing:} \todo{To be written}{Przemek}

\subsection{Operation of History-driven Coalescing}: \sys determines $C$ for a given operation based on the number of tasks completed, $S$, during the current operational period. \sys begins execution in a coalesced task with $C$ initialized to half of the number of tasks executed during the previous operating period and $S$ initialized to $0$. The system executes $C-1$ coalesced tasks followed by one committing task. After each commit, \sys assigns $C$ to half of $S$. If more tasks complete in an operating period, \sys attempts to coalesce more tasks. If fewer tasks
complete, \sys attempts to coalesce fewer tasks. If no committing task
completes due to a power failure, \sys updates $C$ on reboot to half of $S$
(the number of tasks completed before the power failure) and continues.

\noindent\textbf{Task Coalescing Illustration:}
Figure~\ref{fig:coalescing_example_states} shows a sequence of three tasks
executing in \sys. At left, there is no coalescing and tasks incur the cost of
commit at each transition. The center shows the same tasks executing as
coalesced groups of two tasks $C=2$, avoiding one commit.  At right, \sys
coalesces three tasks ($C=3$) avoiding two commits.
Figure~\ref{fig:coalescing_example_flowchart} shows how \sys adapts the value
of $C$ based on $S$. \sys executes tasks in the first operating period, with
$C$ equal to half the number of tasks completed in the prior execution period.
In the example, $C$ starts at eight. After completing eight tasks, \sys commits
and reduces $C$ by half to four.  \sys completes two of four tasks and power
fails.  During the operating period, \sys executed a total of ten tasks (i.e.,
$S=10$) and when \sys restarts, $C$ is assigned to five (i.e., half of $S$).
\sys completes the five coalesced tasks $S=5$ and after committing, assigns $C$
to half of $S$, which is three (after rounding up). \sys coalesces three tasks
and again commits.  At this point, $S=8$; \sys assigns $C=4$ and execution
continues.

@@

\begin{figure}
	\centering
	\includegraphics[width=\columnwidth]{figures/TskDist.pdf}
	\caption{Result of benchmark applications (Section~ref{}) task size measurement, measured in the number of instruction cycles. We observe very large variations of task sizes per each benchmark. \todo{Describe X axis, no capitals in titles - except first word; markers in addition to colors per app; remove 4.7}{Amjad}}
\label{fig:task_profiling}
\end{figure}

>>>>

\textbf{Process of Task Coalescing.} \sys, as the first task-based execution model, dynamically coalesces consecutively executing tasks at run time. When $X$ tasks are coalesced, the $X-1$ tasks do not commit their updates to protected memory locations when reaching a \transition statement. Instead, execution immediately continues from the beginning of the $X$th task. At the end of the $X$th coalesced task, \sys commits the pages that were updated by any of the $X$ coalescing tasks. If consecutive tasks access the same pages of data, the pages accessed by both tasks only need to be committed once. \sys repeatedly executes $X-1$ coalesced tasks, followed by a committing task.

\textbf{Coalescing Speed/Wasted Effort Trade-off.} \sys can coalesce an arbitrary number of consecutive tasks. As more tasks coalesce, their collective commit overhead amortizes better, potentially improving performance. However, as more tasks coalesce, the amount of potentially {\em wasted work} also increases. If power fails during a long sequence of coalesced tasks the lack of an intervening commit requires execution to restart from the {\em first} task in the sequence, failing to preserve the progress of any of the tasks in this long sequence. The challenge to coalescing tasks is determining how many tasks to coalesce before committing. During execution, an executing task is either a {\em coalesced} task, which does not commit its state before transitioning or a {\em committing} task, which does commit its state before transitioning.

\textbf{Task Coalescing Strategies.} The simplest strategy will coalesce/de-coalesce fixed number of task at every successful/unsuccessful commit operation, where task coalescing factor increases by $C_U$ at each successful virtual task execution and decreases by $C_D$, otherwise. There are naturally infinite possible strategies for coalescing, i.e. infinite amount of $(C_U,C_D)$ pairs. The caveat of predefined $(C_U,C_D)$ pair is inability to adapt to changing energy availability conditions. A \emph{good} coalescing strategy must moderates the risk of coalescing and capitalize on its benefit by dynamically, \emph{adaptively} determining a coalescing factors $(C_U,C_D)$ that combines many tasks without suffering an excess of wasted work. Therefore we propose to consider two fundamental coalescing strategies for \sys:

\begin{enumerate}
	\item \textbf{Power Interrupt History-driven Coalescing:} ; and 
	\item \textbf{Power Interrupt History-driven Coalescing with Weights:} where coalescing strategy takes into account \emph{expected size} of each task in the application to pick the best coalescing factor.
\end{enumerate}

\subsection{Operation of Power Interrupt History-driven Coalescing.} \sys determines $C$ for a given operation based on the number of tasks completed, $S$, during the current operational period. \sys begins execution in a coalesced task with $C$ initialized to half of the number of tasks executed during the previous operating period and $S$ initialized to $0$. The system executes $C-1$ coalesced tasks followed by one committing task. After each commit, \sys assigns $C$ to half of $S$. If more tasks complete in an operating period, \sys attempts to coalesce more tasks. If fewer tasks complete, \sys attempts to coalesce fewer tasks. If no committing task completes due to a power failure, \sys updates $C$ on reboot to half of $S$ (the number of tasks completed before the power failure) and continues.

\textbf{Task Coalescing Illustration.} Figure~\ref{fig:coalescing_example_states} shows a sequence of three tasks executing in \sys. At left, there is no coalescing and tasks incur the cost of commit at each transition. The center shows the same tasks executing as coalesced groups of two tasks $C=2$, avoiding one commit.  At right, \sys
coalesces three tasks ($C=3$) avoiding two commits. Figure~\ref{fig:coalescing_example_flowchart} shows how \sys adapts the value
of $C$ based on $S$. \sys executes tasks in the first operating period, with
$C$ equal to half the number of tasks completed in the prior execution period.
In the example, $C$ starts at eight. After completing eight tasks, \sys commits
and reduces $C$ by half to four.  \sys completes two of four tasks and power
fails.  During the operating period, \sys executed a total of ten tasks (i.e.,
$S=10$) and when \sys restarts, $C$ is assigned to five (i.e., half of $S$).
\sys completes the five coalesced tasks $S=5$ and after committing, assigns $C$
to half of $S$, which is three (after rounding up). \sys coalesces three tasks
and again commits.  At this point, $S=8$; \sys assigns $C=4$ and execution
continues.

@@

\begin{figure}
	\centering
	\includegraphics[width=\columnwidth]{figures/TskDist.pdf}
	\caption{Result of benchmark applications (Section~\ref{}) task size measurement, measured in the number of instruction cycles. We observe very large variations of task sizes per each benchmark. \todo{Describe X axis, no capitals in titles - except first word; markers in addition to colors per app; remove 4.7}{Amjad}}
\label{fig:task_profiling}
\end{figure}
